Changes in accessibility to emergency and community food services during COVID-19 and implications for low income populations in Hamilton, Ontario

### 2021-08-27 Accessibility-Foodbanks-Hamilton/Accessibility-Foodbanks-Hamilton.tex
Revised table to show reduction of population serviced during covid, and proofread paper

The impacts of food insecurity during the pandemic are alarming, since
diet-related diseases, such as obesity, heart-disease, and diabetes,
were already critical public health concerns in Canada prior to COVID-19
(Boucher et al., 2017). While emergency food services are not
necessarily a stable solution to food insecurity and in fact may
encourage a retrenchment of neoliberal policy (Wakefield et al., 2013),
at least they can be argued to provide a resource of last instance to
households in precarious situations (Bazerghi et al., 2016). As a
mid-size city grappling with deindustrialization, Hamilton exhibits high
rates of poverty and food bank use. As recently as 2019, the Hamilton
Hunger Report (HFS, 2019) noted that food banks in the city recorded the
highest number of visitors in the past 29 years; a rate of increase
greater than population growth. Most troubling, approximately 40\% of
all visitors were children.

@@

The study of place-based geographic accessibility is concerned with
capturing the potential to reach destinations of value using the
transportation network (Páez et al., 2012). Indeed, the Government of
Canada's recent Food Policy (Agriculture and Canada, 2019) has made
``access'' to healthy food a priority for Canadian
communities{[}\^{}2{]} and previous survey research has suggested that
such accessibility plays a key role in user satisfaction with food bank
service delivery (Holmes et al., 2018). However, as with research into
the prevalence of food deserts, accessibility to food banks is unlikely
to be evenly distributed, and variation throughout a city can be
expected due to transportation network characteristics and the spatial
distribution of service locations and the population they are meant to
serve. Furthermore, policy responses to the COVID-19 pandemic likely
have added to the distress of vulnerable households. Non-pharmaceutical
interventions during the pandemic involving restrictions in mobility
have increased the friction of travel, in particular by transit on which
low income populations are more reliant (e.g., DeWeese et al., 2020). At
the same time, the pandemic has created additional stress for the
operators of food banks through disruptions in the supply chain (e.g.,
McKay et al., 2021) as well as concerns surrounding the delivery of
service in safe conditions and possible cancellation of food service
programs.

For this study, we aim to look at how the landscape of food banks and
related services (e.g.~low-cost or free meal service providers)
available in Hamilton, Ontario, changed during the pandemic. Did the
number of open food bank services diminish? If so, what was the
accessibility to food banks before the pandemic from the perspective of
low income households, and how has it changed during the pandemic with
respect to geographic access and congestion at remaining sites? And
finally, who are most likely to have been impacted by changes in the
accessibility landscape? This paper first looks at the distribution of
food banks and related services before and during the pandemic. Then, we
use the balanced floating catchment area approach of Paez et al. (2019)
to investigate the accessibility situation. For this, we adopt a fully
disaggregated approach based on parcel-level data. Socio-economic and
demographic data are drawn from the latest Census of Canada (2016),
whereas travel information is from the most recent regional travel
survey from 2016. This paper follows reproducible research
recommendations (see Brunsdon and Comber, 2020), and the research was
conducted using open source tools for transportation analysis (Lovelace,
2021). The code and data necessary to reproduce the analysis are
available in a public repository (see Supplemental Material).

\hypertarget{food-insecurity-and-food-bank-use-in-canada}{%
\section{Food Insecurity and Food Bank Use in
Canada}\label{food-insecurity-and-food-bank-use-in-canada}}

@@

Nationally, it has previously been found that food insecurity impacts
approximately 12.3\% of Canadian households (Tarasuk et al., 2014) using
data from the 2011-2012 CCHS. Using the same data, Tarasuk et al. (2019)
found higher odds of food insecurity amongst households relying on
social assistance, those without a university degree or with children
under the age of 18, and individuals that lived alone, renters, and
those identifying as Aboriginal. While surveys revealed that only 20 to
30 percent of those experiencing food insecurity were found to frequent
food banks in Canada (Tarasuk et al., 2014), pre-pandemic research from
Ottawa (Enns et al., 2020) and Vancouver (Black and Seto, 2020) suggests
that long-term users tend to be older, have health or mobility
challenges, live in large households, and are less likely to have
employment income. In terms of geography, previous research conducted at
the provincial scale using data from the 2011-2012 CCHS found that the
prevalence of food insecurity ranged across the country from 11.8\% of
households in Ontario to 41\% of households in Nunavut (Tarasuk et al.,
2019).

@@

In the United States, economic stress caused by the COVID-19 pandemic,
rising unemployment rates, and changes in poverty levels have disrupted
the food environment and led to higher rates of food insecurity (Niles
et al., 2020). Pandemic-related food security studies in the US have
found a substantial increase in households experiencing food insecurity
for the first time, and also in households experiencing more severe food
insecurity than before (Niles et al., 2020; Wolfson and Leung, 2020).
Considering the negative mental and physical health effects, increases
in food insecurity rates due to the onset of COVID-19, signal a change
in the food environment with potential damages to population health
during the course of the pandemic and beyond (Niles et al., 2020).

Canada has also seen an increase in the number of households living in
food insecurity due to the COVID-19 pandemic. Survey results from
Statistics Canada from May of 2020 suggest that 14.7\% of the population
was living in food insecurity in the past 30 days, up from 10.5\% in
2017-2018 (Statistics Canada, 2020a). Recent data from FBC (FBC, 2020)
showed that 52\% of member food banks reported an increase in usage in
March of 2020 when initial lockdown restrictions were put in place
across much of the country. The pandemic also created significant
staffing issues with 42\% of food banks reporting a reduction in
volunteers. However, 53\% of food banks later reported a decrease in use
into the summer of 2020 which FBC members attributed to emergency
financial support programs from the federal government. Nevertheless,
some of these benefit programs were temporary. Although more recent
statistics on food bank use in Hamilton in 2020 and 2021 are not yet
available, data from the Daily Bread Food Bank (DBFB, 2020) in
neighbouring Toronto for August 2020 shows visits climbing 51\%
year-over-year in that city, which suggests that many households in
Hamilton are likely to turn to food bank services to meet their needs.

@@

\[
w_{ij}^\text{st} = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
\]

@@

\[
w_{ji}^\text{st} = \frac{w_{ji}}{\sum_{j=1}^Jw_{ji}}
\]

@@

\[
\sum_{i=1}^nw_{ij}^\text{st}=1
\]

@@

\[
\sum_{j=1}^Jw_{ji}^\text{st}=1
\]

@@

\[
\sum_{i=1}^nP_iw_{ij}^\text{st}=P_i
\]

@@

\[
\sum_{j=1}^JL_jw_{ji}^\text{st}=L_j
\]

@@

Data have been prepared for sharing as a data package available in the
Supplemental Materials. The contents of the data package are described
next.

@@

\[
w_{ij}=w_{ji}=
\begin{cases}
1 & \text{ if } ett_{ij}\le \delta\\
0 & \text{ otherwise}
\end{cases}
\]

@@

Using the 15-minute threshold, we find that the system-wide
accessibility (interpreted as a provider-to-population ratio) was 0.078
(food banks/service locations per low income household in the city)
before COVID-19, but declined to 0.048 during the pandemic. It is
striking that although almost 76\% of food facilities remained in
operation during the pandemic, there was a loss of accessibility greater
than 39\%, suggesting the location of emergency and community food
services plays an important role in serving those in need.

@@

For the more suburban clusters of zones, the decrease in accessibility
is derived from the closure of locations throughout the city reachable
by car. In the cluster of central suburban zones for example, low-income
households in the outer ring of zones that exhibit medium to high
decreases in accessibility within this cluster appear to be largely
auto-dependent in their tripmaking, which each each exhibiting between
85-100\% of their modal split for car trips. This results in the parcels
within these zones having a large number of potentially accessible
locations in the travel time matrix. But by extension, the change in
accessibility over the pre- and during-COVID-19 time periods is affected
not only by the closure of service locations proximate to the zones, but
also the locations in the central city. The zone with the greatest
decrease in accessibility within this cluster (-0.0009) has a high rate
of car trips and connects to the most food bank locations in total as
well as those that stayed open or closed.

In the cluster to the south-west, the decrease in accessibility is
predominately driven by the closure of a high level-of-service Community
Meals provider. However, like the more central suburban zones,
low-income households within this cluster are also between 90\% to 100\%
auto-dependent in their tripmaking in the TTS. The story is similar for
the zone located in the north-west that exhibits the greatest decrease
in accessibility. Here, low-income households responding to the TTS
conducted 100\% of their trips by car and, as a result, dwellings within
this zone have access to the second-highest number of food bank and
service locations within 15 minutes. However, this also means zonal
accessibilities are greatly affected by the number of closures
throughout the city. Finally, in the city's north end and north-east
zones, low-income households exhibit a mixture of tripmaking behaviour
in the TTS. Households in some zones take transit more often and one
zone in particular has 100\% of its trips by walking. For these zones,
the decrease in accessibility tends to be a product of the closure of
several inner-city food bank and service locations reachable by multiple
modes.

Just as the effects of the closures appear to have been uneven in space,
they also seem to have had different impacts on various population
segments. Using data on low income individuals by age drawn from the
TTS, Table \ref{tab:accessibility-by-age} shows the estimated number of
people in each age group by their level of accessibility before and
during the pandemic. Here, it is important to note that the quartiles
are relative: people in the top 25\% of accessibility still have lower
accessibility during the pandemic than before. In reality, every
population group is worse off during the pandemic in terms of their
accessibility to food banks and services in the City of Hamilton.
However, some age groups were affected more. As seen in the table, the
distribution within quartiles for adults changed only slightly, and in
fact there are fewer adults in the bottom three quartiles as some found
themselves in the (deteriorated) top quartile of accessibility during
the pandemic. The story is different for seniors, a greater number of
whom are now in the third quartile, due to a combination of seniors
facing a worse accessibility situation, and just a few moving from the
bottom quartile (pre-pandemic) to the third during the pandemic. Those
aged 18 and less appear to have fared worse, and the largest change is
in the number of children who were in the third quartile before the
pandemic and found themselves in the bottom accessibility class during
the pandemic. These results suggest that, through a combination of the
typical modes of transportation of lower income households and the
spatial distribution of the population, the food bank/service locations
closed had a differential impact that more greatly affected the youngest
among the population in low income households.

@@

\caption{\label{tab:table-accessibility-by-age-group}\label{tab:accessibility-by-age}Population at each accessibility level by age group among members households with incomes less than CAD40,000.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Pre-COVID-19} & \multicolumn{3}{c}{During COVID-19} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
Accessibility Quartile & Children (age $\le$ 18) & Adults (19-64) & Seniors (age $\ge$ 65) & Children (age $\le$ 18) & Adults (19-64) & Seniors (age $\ge$ 65)\\
\midrule
\cellcolor{gray!6}{Top 25\%} & \cellcolor{gray!6}{2,048} & \cellcolor{gray!6}{7,031} & \cellcolor{gray!6}{3,632} & \cellcolor{gray!6}{1,956 (-4.49\%)} & \cellcolor{gray!6}{6,138 (-12.7\%)} & \cellcolor{gray!6}{3,338 (-8.09\%)}\\
Second 25\% & 2,190 & 9,631 & 5,023 & 2,365 (7.99\%) & 10,833 (12.48\%) & 5,400 (7.51\%)\\
\cellcolor{gray!6}{Third 25\%} & \cellcolor{gray!6}{3,239} & \cellcolor{gray!6}{12,694} & \cellcolor{gray!6}{6,875} & \cellcolor{gray!6}{3,082 (-4.85\%)} & \cellcolor{gray!6}{12,612 (-0.65\%)} & \cellcolor{gray!6}{7,064 (2.75\%)}\\
Bottom 25\% & 3,711 & 13,220 & 7,648 & 3,038 (-18.14\%) & 11,811 (-10.66\%) & 6,308 (-17.52\%)\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Population values have been rounded.}\\
\multicolumn{7}{l}{\rule{0pt}{1em}The values in brackets for population during COVID-19 are the changes from before the pandemic.}\\
\end{tabular}}
\end{table}

@@

Previous work has noted the important role of geography alongside other
socio-economic and demographic indicators in household access to healthy
food. The present papers is, to the best of our knowledge, among the
first studies to focus on the geographic component of accessibility to
emergency food services (Allen and Farber, 2021). Using the balanced
floating catchment area method to account for population demand and
congestion effects at service points, we estimated multi-modal
accessibility to emergency and community food service locations for
low-income households. The weighting of travel time estimates by the
modal split in different zonal geographies tailors the results to
patterns of travel behaviour captured in the regional travel survey.
Moreover, our parcel-level analysis presents a disaggregate approach to
estimating accessibility based on the locations of residential parcels
and dwellings. Beyond accounting for the inflation of demand that occurs
in traditional floating catchment area methods, our application of the
balanced floating catchment area approach offers a novel analysis of
accessibility to emergency and communal food services that is sensitive
to the locations of low-income households, information on their age
distributions and typical trip-making behaviour, the locations of
services, their operations over time, and the characteristics of the
city's multi-modal transportation network.

@@

It is important to note that the degree to which our low-income cut-off
of CAD\$40,000 reflects food insecurity in the different zones of the
study area is not known. We also consider all emergency and communal
food service locations equally. Information on differential capacities
at food provider locations is currently not collected, and given the
closure of facilities was not possible to obtain. Data on services, such
as the number of meals served, could be used to refine the analysis in
the future. Moreover, while the travel survey allows us to model
multi-modal accessibilities that align with travel behaviour observed in
the travel survey and capture differences in accessibility by age
categories, the use of travel survey data for modelling food insecurity
also has its limitations. Research into the population weighting methods
used in the TTS note that the survey may under-count the lowest- and
highest-income households in the survey study region, although the
magnitude of this under-counting is unknown, and approximately 20\% of
respondents to the 2016 survey did not report their income (Rose, 2018).
In that regard, the modal splits of low-income households observed
through the TTS data may not accurately reflect the travel behaviours of
food insecure populations, and our estimates might in fact be somewhat
conservative if those households who do not report income rely more on
walking and/or transit for their mobility needs.

@@

@@

\leavevmode\hypertarget{ref-niles2020early}{}%
Niles, M.T., Bertmann, F., Belarmino, E.H., Wentworth, T., Biehl, E.,
Neff, R., 2020. The early food insecurity impacts of COVID-19. Nutrients
12, 2096.

@@

\leavevmode\hypertarget{ref-wolfson2020food}{}%
Wolfson, J.A., Leung, C.W., 2020. Food insecurity and COVID-19:
Disparities in early effects for US adults. Nutrients 12, 1648.

>>>>

The impacts of food insecurity during the pandemic are alarming, since
diet-related diseases, such as obesity, heart-disease, and diabetes,
were already critical public health concerns in Canada prior to COVID-19
(Boucher et al., 2017). While emergency food services are not
necessarily a stable solution to food insecurity and in fact may
encourage a retrenchment of neoliberal policy (Wakefield et al., 2013),
in reality provide a resource of last instance to households in
precarious situations (Bazerghi et al., 2016). As a mid-size city
grappling with deindustrialization, Hamilton exhibits high rates of
poverty and use of emergency food services. As recently as 2019, the
Hamilton Hunger Report (HFS, 2019) noted that food banks in the city
recorded the highest number of visitors in the past 29 years; a rate of
increase greater than population growth. Most troubling, approximately
40\% of all visitors were children.

@@

The study of place-based geographic accessibility is concerned with
capturing the potential to reach destinations of value using the
transportation network (Páez et al., 2012). Indeed, the Government of
Canada's recent Food Policy (Agriculture and Canada, 2019) has made
``access'' to healthy food a priority for Canadian communities and
previous research suggests that such accessibility plays a key role in
user satisfaction with food bank service delivery (Holmes et al., 2018).
However, as with research into the prevalence of food deserts,
accessibility to food banks is unlikely to be evenly distributed, and
variation throughout a city can be expected due to transportation
network characteristics and the spatial distribution of service
locations and the population they are meant to serve. Furthermore,
policy responses to the COVID-19 pandemic likely have added to the
distress of vulnerable households. Non-pharmaceutical interventions
during the pandemic involving restrictions in mobility have increased
the friction of travel, in particular by transit on which low income
populations are more reliant (e.g., DeWeese et al., 2020). At the same
time, the pandemic has created additional stress for the operators of
food banks through disruptions in the supply chain (e.g., McKay et al.,
2021) as well as concerns surrounding the delivery of service in safe
conditions and possible cancellation of food service programs.

For this study, we aim to look at how the landscape of emergency food
and related services (e.g.~low-cost or free meal service providers)
available in Hamilton, Ontario, changed during the pandemic. Did the
number of open services diminish? If so, what was the accessibility to
emergency and community food services before the pandemic from the
perspective of low income households, and how has it changed during the
pandemic with respect to geographic access and congestion at remaining
sites? And finally, who are most likely to have been impacted by changes
in the accessibility landscape? This paper first looks at the
distribution of emergency and community food services before and during
the pandemic. Then, we use the balanced floating catchment area approach
of Paez et al. (2019) to investigate the accessibility situation. For
this, we adopt a fully disaggregated approach based on parcel-level
data. Socio-economic and demographic data are drawn from the latest
Census of Canada (2016), whereas travel information is from the most
recent regional travel survey from 2016. This paper follows reproducible
research recommendations (see Brunsdon and Comber, 2020), and the
research was conducted using open source tools for transportation
analysis (Lovelace, 2021). The data and code needed to reproduce the
analysis are available in a public repository (see Supplemental
Material).

\hypertarget{food-insecurity-and-emergency-and-communal-food-services-in-canada}{%
\section{Food Insecurity and Emergency and Communal Food Services in
Canada}\label{food-insecurity-and-emergency-and-communal-food-services-in-canada}}

@@

Nationally, analysis of the 2011-2012 CCHS has previously revealed that
food insecurity impacts approximately 12.3\% of Canadian households
(Tarasuk et al., 2014). Using the same data, Tarasuk et al. (2019) found
higher odds of food insecurity amongst households relying on social
assistance, those without a university degree or with children under the
age of 18, and individuals that lived alone, renters, and those
identifying as Aboriginal. While surveys revealed that only 20 to 30
percent of those experiencing food insecurity were found to frequent
food banks in Canada (Tarasuk et al., 2014), pre-pandemic research from
Ottawa (Enns et al., 2020) and Vancouver (Black and Seto, 2020) suggests
that long-term users tend to be older, have health or mobility
challenges, live in large households, and are less likely to have
employment income. In terms of geography, previous research conducted at
the provincial scale using data from the 2011-2012 CCHS found that the
prevalence of food insecurity ranged across the country from 11.8\% of
households in Ontario to 41\% of households in Nunavut (Tarasuk et al.,
2019).

@@

Canada also saw an increase in the number of households living in food
insecurity due to the COVID-19 pandemic. Survey results from Statistics
Canada from May of 2020 suggest that 14.7\% of the population was living
in food insecurity in the past 30 days, up from 10.5\% in 2017-2018
(Statistics Canada, 2020a). Recent data from FBC (FBC, 2020) showed that
52\% of member food banks reported an increase in usage in March of 2020
when initial lockdown restrictions were put in place across much of the
country. The pandemic also created significant staffing issues with 42\%
of food banks reporting a reduction in volunteers. However, 53\% of food
banks later reported a decrease in use into the summer of 2020 which FBC
members attributed to emergency financial support programs from the
federal government. Nevertheless, some of these benefit programs were
temporary. Although more recent statistics on food bank use in Hamilton
in 2020 and 2021 are not yet available, data from the Daily Bread Food
Bank (DBFB, 2020) in neighbouring Toronto for August 2020 shows visits
climbing 51\% year-over-year in that city, which suggests that many
households in Hamilton are likely to turn to food bank services to meet
their needs.

@@

\[
w_{ij}^i = \frac{w_{ij}}{\sum_{i=1}^nw_{ij}}
\]

@@

\[
w_{ij}^j = \frac{w_{ij}}{\sum_{j=1}^Jw_{ij}}
\]

@@

\[
\sum_{i=1}^nw_{ij}^i=1
\]

@@

\[
\sum_{j=1}^Jw_{ij}^j=1
\]

@@

\[
\sum_{i=1}^nP_iw_{ij}^i=P_i
\]

@@

\[
\sum_{j=1}^JL_jw_{ij}^j=L_j
\]

@@

Data have been prepared for sharing in the form of an open data product
(see Arribas-Bel et al., 2021) available in the Supplemental Materials.
The contents of the data package are described next.

@@

\[
w_{ij}=
\begin{cases}
1 & \text{ if } ett_{ij}\le \delta\\
0 & \text{ otherwise}
\end{cases}
\]

@@

Using the 15-minute threshold, we find that the system-wide
accessibility was 0.078 (food banks/service locations per low income
household in the city) before COVID-19, but declined to 0.048 during the
pandemic. It is striking that although almost 76\% of facilities
remained in operation during the pandemic, there was a loss of
accessibility greater than 39\%, suggesting the location of emergency
and community food services plays an important role in serving those in
need.

@@

For the more suburban clusters of zones, the decrease in accessibility
is derived from the closure of locations throughout the city reachable
by car. In the cluster of central suburban zones for example, low-income
households in the outer ring of zones that exhibit medium to high
decreases in accessibility within this cluster appear to be largely
auto-dependent in their tripmaking, which each each exhibiting between
85-100\% of their modal split for car trips. This results in the parcels
within these zones having a large number of potentially accessible
locations in the travel time matrix. But by extension, the change in
accessibility over the pre- and during-COVID-19 time periods is affected
not only by the closure of service locations proximate to the zones, but
also the locations in the central city. The zone with the greatest
decrease in accessibility within this cluster (-0.0009) has a high rate
of car trips and connects to the most facility locations in total as
well as those that stayed open or closed.

In the cluster to the south-west, the decrease in accessibility is
predominately driven by the closure of a high level-of-service Community
Meals provider. However, like the more central suburban zones,
low-income households within this cluster are also between 90\% to 100\%
auto-dependent in their trip-making in the TTS. The story is similar for
the zone located in the north-west that exhibits the greatest decrease
in accessibility. Here, low-income households responding to the TTS
conducted 100\% of their trips by car and, as a result, dwellings within
this zone have access to the second-highest number of food bank and
service locations within 15 minutes. However, this also means zonal
accessibilities are greatly affected by the number of closures
throughout the city. Finally, in the city's north end and north-east
zones, low-income households exhibit a mixture of tripmaking behaviour
in the TTS. Households in some zones take transit more often and one
zone in particular has 100\% of its trips by walking. For these zones,
the decrease in accessibility tends to be a product of the closure of
several inner-city food bank and service locations reachable by multiple
modes.

Just as the effects of the closures appear to have been uneven in space,
they also seem to have had different impacts on various population
segments. Using data on low income individuals by age drawn from the
TTS, Table \ref{tab:accessibility-by-age} shows the estimated number of
people in each age group by their level of accessibility before and
during the pandemic. Here, it is important to note that the quartiles
are relative: people in the top 25\% of accessibility still have lower
accessibility during the pandemic than before. In reality, every
population group is worse off during the pandemic in terms of their
accessibility to food banks and services in the City of Hamilton.
However, some age groups were affected more. As seen in the table, the
distribution within quartiles for adults changed only slightly, and in
fact there are fewer adults in the bottom three quartiles as some found
themselves in the (deteriorated) top quartile of accessibility during
the pandemic. The story is different for seniors, a greater number of
whom are now in the third quartile, due to a combination of seniors
facing a worse accessibility situation, and just a few moving from the
bottom quartile (pre-pandemic) to the third during the pandemic. Those
aged 18 and less appear to have fared worse, and the largest change is
in the number of children who were in the third quartile before the
pandemic and found themselves in the bottom accessibility class during
the pandemic. In fact, when we compare the total population serviced
before and during the pandemic, we see that a large number of children,
adults, and seniors were no longer in the catchment areas of service
locations (last row of Table \ref{tab:accessibility-by-age}). It is
remarkable that despite the loss of population serviced, accessibility
still declined for those still within the catchment regions of these
services.

These results suggest that, through a combination of the typical modes
of transportation of lower income households and the spatial
distribution of the population, the food bank/service locations closed
had a differential impact that more greatly affected the youngest among
the population in low income households.

@@

\caption{\label{tab:table-accessibility-by-age-group}\label{tab:accessibility-by-age}Population at each accessibility level by age group among members households with incomes less than CAD40,000.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Pre-COVID-19} & \multicolumn{3}{c}{During COVID-19} \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
Accessibility Quartile & Children (age $\le$ 18) & Adults (19-64) & Seniors (age $\ge$ 65) & Children (age $\le$ 18) & Adults (19-64) & Seniors (age $\ge$ 65)\\
\midrule
\cellcolor{gray!6}{Top 25\%} & \cellcolor{gray!6}{2,048} & \cellcolor{gray!6}{7,031} & \cellcolor{gray!6}{3,632} & \cellcolor{gray!6}{1,956 (-4.49\%)} & \cellcolor{gray!6}{6,138 (-12.7\%)} & \cellcolor{gray!6}{3,338 (-8.09\%)}\\
Second 25\% & 2,190 & 9,631 & 5,023 & 2,365 (7.99\%) & 10,833 (12.48\%) & 5,400 (7.51\%)\\
\cellcolor{gray!6}{Third 25\%} & \cellcolor{gray!6}{3,239} & \cellcolor{gray!6}{12,694} & \cellcolor{gray!6}{6,875} & \cellcolor{gray!6}{3,082 (-4.85\%)} & \cellcolor{gray!6}{12,612 (-0.65\%)} & \cellcolor{gray!6}{7,064 (2.75\%)}\\
Bottom 25\% & 3,711 & 13,220 & 7,648 & 3,038 (-18.14\%) & 11,811 (-10.66\%) & 6,308 (-17.52\%)\\
\midrule
\cellcolor{gray!6}{Total population} & \cellcolor{gray!6}{11,189} & \cellcolor{gray!6}{42,576} & \cellcolor{gray!6}{23,178} & \cellcolor{gray!6}{10,441 (-6.69\%)} & \cellcolor{gray!6}{41,395 (-2.77\%)} & \cellcolor{gray!6}{22,110 (-4.61\%)}\\
\bottomrule
\multicolumn{7}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{7}{l}{\rule{0pt}{1em}Population values have been rounded.}\\
\multicolumn{7}{l}{\rule{0pt}{1em}The values in brackets for population during COVID-19 are the changes from before the pandemic.}\\
\end{tabular}}
\end{table}

@@

Previous work has noted the important role of geography alongside other
socio-economic and demographic indicators in household access to healthy
food. The present papers is, to the best of our knowledge, among the
first studies to focus on the geographic component of accessibility to
emergency food services (Allen and Farber, 2021). Using the balanced
floating catchment area method to account for population demand and
congestion effects at service points, we estimated multi-modal
accessibility to emergency and community food service locations for
low-income households. The weighting of travel time estimates by the
modal split in different zonal geographies tailors the results to
patterns of travel behaviour captured in the regional travel survey.
Moreover, our parcel-level analysis presents a disaggregate approach to
estimating accessibility based on the locations of residential parcels
and dwellings. Beyond accounting for the inflation of demand that occurs
in traditional floating catchment area methods, our application of the
balanced floating catchment area approach offers a novel analysis of
accessibility to emergency and community food services that is sensitive
to the locations of low-income households, information on their age
distributions and typical trip-making behaviour, the locations of
services, their operations over time, and the characteristics of the
city's multi-modal transportation network.

@@

It is important to note that the degree to which our low-income cut-off
of CAD\$40,000 reflects food insecurity in the different zones of the
study area is not known. We also consider all emergency and community
food service locations equally. Information on differential capacities
at food provider locations is currently not collected, and given the
closure of facilities was not possible to obtain. Data on services, such
as the number of meals served, could be used to refine the analysis in
the future. Moreover, while the travel survey allows us to model
multi-modal accessibilities that align with travel behaviour observed in
the travel survey and capture differences in accessibility by age
categories, the use of travel survey data for modelling food insecurity
also has its limitations. Research into the population weighting methods
used in the TTS note that the survey may under-count the lowest- and
highest-income households in the survey study region, although the
magnitude of this under-counting is unknown, and approximately 20\% of
respondents to the 2016 survey did not report their income (Rose, 2018).
In that regard, the modal splits of low-income households observed
through the TTS data may not accurately reflect the travel behaviours of
food insecure populations, and our estimates might in fact be somewhat
conservative if those households who do not report income rely more on
walking and/or transit for their mobility needs.

@@

\leavevmode\hypertarget{ref-arribas2021open}{}%
Arribas-Bel, D., Green, M., Rowe, F., Singleton, A., 2021. Open data
products: A framework for creating valuable analysis-ready data. Journal
of Geographical Systems.
doi:\href{https://doi.org/10.1007/s10109-021-00361-7}{10.1007/s10109-021-00361-7}

@@

@@
